---
title: "COMP4441_FinalProject"
author: "Adi Bose, Daniel D'Aquila, Grant Sheffer"
date: "8/9/2021"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(ggpubr)
library(boot)
library(lawstat)
library(ggplot2)
library(lattice)
library(caret)
```

Weekly data, IndexNum indicates which week it is of each year. 

Loading Data into R Studio 
```{r}
avocadoData <- read.csv("avocado.csv")
avocadoData <- filter(avocadoData, year != 2018)

avocadoData2018 <- read.csv("2018avocadoClean.csv")

sapply(avocadoData, class)
sapply(avocadoData2018, class)

avocadoData$Date <- as.Date(avocadoData$Date,format="%Y-%m-%d")
avocadoData2018$Date <- as.Date(avocadoData2018$Date,format="%m/%d/%y")

colnames(avocadoData)[colnames(avocadoData) == "X"] <- "IndexNum"
colnames(avocadoData2018)[colnames(avocadoData2018) == "Index"] <- "IndexNum"

avocadoData$regionFactor <- factor(avocadoData$region)
avocadoData2018$regionFactor <- factor(avocadoData2018$region)

avocadoData$typeFactor <- factor(avocadoData$type)
avocadoData2018$typeFactor <- factor(avocadoData2018$type)

avocadoData$DateNumeric <- as.numeric(avocadoData$Date)
avocadoData2018$DateNumeric <- as.numeric(avocadoData2018$Date)

sapply(avocadoData, class)
sapply(avocadoData2018, class)

```


Data filtering by regions
```{r}
# Cleaning the data by regions [removing the 'state' entries]

table(avocadoData$region)

regions = c("Midsouth", "Northeast", "Plains", "SouthCentral", "Southeast", "West", "California", "GreatLakes")
AvocadoRegions <- filter(avocadoData, region %in% regions)

AvocadoRegionsConven <- filter(AvocadoRegions, type == "conventional")

AvocadoRegionsOrg <- filter(AvocadoRegions, type == "organic")

```


Experimenting with different plots
```{r}
### Average price across all regions wrt time (2015-early 2018)
plot(AveragePrice~Date, data = AvocadoRegions)


### AvocadoRegions plots
## Date vs Volume [comparing regions]
ggplot(AvocadoRegions, aes(Date, Total.Volume)) + geom_point(aes(color = factor(regionFactor))) + theme(legend.position = "bottom")

## Date vs Average Price [comparing regions]
ggplot(AvocadoRegions, aes(Date, AveragePrice)) + geom_point(aes(color = factor(regionFactor))) + theme(legend.position = "bottom")

## Date vs Volume [comparing type]
ggplot(AvocadoRegions, aes(Date, Total.Volume)) + geom_point(aes(color = factor(type))) + theme(legend.position = "bottom")


### Organic 
## Date vs Volume [comparing regions]
ggplot(AvocadoRegionsOrg, aes(Date, Total.Volume)) + geom_point(aes(color = factor(regionFactor))) + theme(legend.position = "bottom")

## Date vs Average Price [comparing regions]
ggplot(AvocadoRegionsOrg, aes(Date, AveragePrice)) + geom_point(aes(color = factor(regionFactor))) + theme(legend.position = "bottom")


### Conventional
## Date vs Volume [comparing regions]
ggplot(AvocadoRegionsConven, aes(Date, Total.Volume)) + geom_point(aes(color = factor(regionFactor))) + theme(legend.position = "bottom")

## Date vs Average Price [comparing regions]
ggplot(AvocadoRegionsConven, aes(Date, AveragePrice)) + geom_point(aes(color = factor(regionFactor))) + theme(legend.position = "bottom")


### Efforts to see which region has the highest demand 
plot(AveragePrice~regionFactor, data = AvocadoRegions) + abline(h = mean(AvocadoRegions$AveragePrice), col = "red")

q <- ggplot(AvocadoRegions, aes(region, Total.Volume)) + 
  geom_point(aes(color = factor(regionFactor)))

q + theme(axis.text.x = element_text(angle = -45, vjust = 1.5, hjust=-0.1),
          legend.position="none")

```


Linear regressions
```{r}
#### Linear Models (Conv + Org)
## Average Price dependent on Date
lm.All.1 <- lm(AveragePrice ~ Date, data = AvocadoRegions)
summary(lm.All.1)

## Average Price dependent on Date AND Total Bags
lm.All.2 <- lm(AveragePrice ~ Date + Total.Bags, data = AvocadoRegions)
summary(lm.All.2)

## Average Price dependent on Date AND Total Bags AND Total Volume (Total Bags showing signs of insignificance)
lm.All.3 <- lm(AveragePrice ~ Date + Total.Bags + Total.Volume, data = AvocadoRegions)
summary(lm.All.3)

## Average Price dependent on Date AND Total Bags AND Total Volume AND Type
lm.All.4 <- lm(AveragePrice ~ Date + Total.Bags + Total.Volume + typeFactor, data = AvocadoRegions)
summary(lm.All.4)

## Average Price dependent on Date AND Total Volume and Type (Total Bags removed due to low significance)
lm.All.5 <- lm(AveragePrice ~ Date + Total.Volume + typeFactor, data = AvocadoRegions)
summary(lm.All.5)

## Average Price dependent on Total Volume and Type (Date removed for comparative purposes)
lm.All.6 <- lm(AveragePrice ~ Total.Volume + typeFactor, data = AvocadoRegions)
summary(lm.All.6)

## Average Price dependent on Total Volume (Type removed for comparative purposes)
lm.All.7 <- lm(AveragePrice ~ Total.Volume, data = AvocadoRegions)
summary(lm.All.7)


#### Linear Models (Conventional)
## Average Price dependent on Date
lm.Conv.1 <- lm(AveragePrice ~ Date, data = AvocadoRegionsConven)
summary(lm.Conv.1)

## Average Price dependent on Date AND Total Bags
lm.Conv.2 <- lm(AveragePrice ~ Date + Total.Bags, data = AvocadoRegionsConven)
summary(lm.Conv.2)

## Average Price dependent on Date AND Total Bags AND Total Volume
lm.Conv.3 <- lm(AveragePrice ~ Date + Total.Bags + Total.Volume, data = AvocadoRegionsConven)
summary(lm.Conv.3)


#### Linear Models (Organic)
## Average Price dependent on Date
lm.Org.1 <- lm(AveragePrice ~ Date, data = AvocadoRegionsOrg)
summary(lm.Org.1)

## Average Price dependent on Date AND Total Bags
lm.Org.2 <- lm(AveragePrice ~ Date + Total.Bags, data = AvocadoRegionsOrg)
summary(lm.Org.2)

## Average Price dependent on Date AND Total Bags AND Total Volume
lm.Org.3 <- lm(AveragePrice ~ Date + Total.Bags + Total.Volume, data = AvocadoRegionsOrg)
summary(lm.Org.3)

```


Polynomial Regressions
```{r}
### General viewing with a stat_smooth() 'regression' added
## Conv + Org (appears to be a 6th order poly fit)
g.All.test <- ggplot(AvocadoRegions, aes(Date, AveragePrice) ) +
                        geom_point() +
                        stat_smooth()
## Conventional (appears to be a 6th order poly fit)
g.Conv.test <- ggplot(AvocadoRegionsConven, aes(Date, AveragePrice) ) +
                        geom_point() +
                        stat_smooth()
## Organic (appears to be a 6th order poly fit)
g.Org.test <- ggplot(AvocadoRegionsOrg, aes(Date, AveragePrice) ) +
                     geom_point() +
                     stat_smooth()


#### Polynomial Models (Conv + Org) ####
## Model 1: 1 IV (Date)
pm.All.1 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE),
                              data = AvocadoRegions)
summary(pm.All.1)

## Model 2: 2 IV's (Date AND Total.Bags)
pm.All.2 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                              poly(Total.Bags, 6, raw = TRUE),
                              data = AvocadoRegions)
summary(pm.All.2)

## Model 3: 3 IV's (Date AND Total.Bags AND Total.Volume)
pm.All.3 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                              poly(Total.Bags, 6, raw = TRUE) +
                              poly(Total.Volume, 6, raw = TRUE),
                              data = AvocadoRegions)
summary(pm.All.3)

## Model 4: 4 IV's (Date AND Total.Bags AND Total.Volume AND Type) (model does not like typeFactor)
# pm.All.4 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
#                               poly(Total.Bags, 6, raw = TRUE) +
#                               poly(Total.Volume, 6, raw = TRUE) +
#                               poly(typeFactor, 6, raw = TRUE),
#                               data = AvocadoRegions)
# summary(pm.All.4)

## Model 5: 2 IV's (Date AND Total.Volume) (removed Type due to incompatibility and Total.Bags due to statistical insignificance)
pm.All.5 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                              poly(Total.Volume, 6, raw = TRUE),
                              data = AvocadoRegions)
summary(pm.All.5)

#### ANOVA (Conv + Org)
## Models 1 and 2 (adding JUST Total.Bags)
anova(pm.All.1, pm.All.2)

## Models 1 and 5 (adding JUST Total.Volume)
anova(pm.All.1, pm.All.5)

## Models 5 and 3 (adding Total.Bags [already having Total.Volume])
anova(pm.All.5, pm.All.3)

## Models 2 and 3 (adding Total.Volume [already having Total.Bags])
anova(pm.All.2, pm.All.3)


#### Polynomial Models (Conventional) ####
## Model 1: 1 IV (Date)
pm.Conv.1 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE),
                               data = AvocadoRegionsConven)
summary(pm.Conv.1)

## Model 2: 2 IV's (Date AND Total.Bags)
pm.Conv.2 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                               poly(Total.Bags, 6, raw = TRUE),
                               data = AvocadoRegionsConven)
summary(pm.Conv.2)

## Model 3: 3 IV's (Date AND Total.Bags AND Total.Volume)
pm.Conv.3 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                               poly(Total.Bags, 6, raw = TRUE) +
                               poly(Total.Volume, 6, raw = TRUE),
                               data = AvocadoRegionsConven)
summary(pm.Conv.3)

## Model 4: 2 IV's (Date AND Total.Volume)
pm.Conv.4 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                               poly(Total.Volume, 6, raw = TRUE),
                               data = AvocadoRegionsConven)
summary(pm.Conv.4)

#### ANOVA (Conventional)
## Models 1 and 2 (adding JUST Total.Bags)
anova(pm.Conv.1, pm.Conv.2)

## Models 1 and 4 (adding JUST Total.Volume)
anova(pm.Conv.1, pm.Conv.4)

## Models 4 and 3 (adding Total.Bags [already having Total.Volume])
anova(pm.Conv.4, pm.Conv.3)

## Models 2 and 3 (adding Total.Volume [already having Total.Bags])
anova(pm.Conv.2, pm.Conv.3)


#### Polynomial Models (Organic) ####
## Model 1: 1 IV (Date)
pm.Org.1 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE),
                              data = AvocadoRegionsOrg)
summary(pm.Org.1)

## Model 2: 2 IV's (Date AND Total.Bags)
pm.Org.2 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                              poly(Total.Bags, 6, raw = TRUE),
                              data = AvocadoRegionsOrg)
summary(pm.Org.2)

## Model 3: 3 IV's (Date AND Total.Bags AND Total.Volume)
pm.Org.3 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                              poly(Total.Bags, 6, raw = TRUE) +
                              poly(Total.Volume, 6, raw = TRUE),
                              data = AvocadoRegionsOrg)
summary(pm.Org.3)

## Model 4: 2 IV's (Date AND Total.Volume)
pm.Org.4 <- lm(AveragePrice ~ poly(DateNumeric, 6, raw = TRUE) +
                              poly(Total.Volume, 6, raw = TRUE),
                              data = AvocadoRegionsOrg)
summary(pm.Org.4)

#### ANOVA (Organic)
## Models 1 and 2 (adding Total.Bags)
anova(pm.All.1, pm.All.2)

## Models 2 and 3 (adding Total.Volume)
anova(pm.All.2, pm.All.3)

## Models 4 and 3 (adding Total.Bags [already having Total.Volume])
anova(pm.Org.4, pm.Org.3)

## Models 2 and 3 (adding Total.Volume [already having Total.Bags])
anova(pm.Org.2, pm.Org.3)
```


Price Predictions
```{r}
#### Predictions for Conv + Organic
## pm.All.3
price.pred.All <- data.frame(predict(pm.All.3, avocadoData2018))
names(price.pred.All) <- c("pred")
price.pred.All$actual <- avocadoData2018$AveragePrice
price.pred.All$diff <- abs(price.pred.All$pred - price.pred.All$actual)
mean(price.pred.All$diff)

#### Predictions for Conventional
## pm.Conv.3
price.pred.Conv <- data.frame(predict(pm.Conv.4, avocadoData2018))
names(price.pred.Conv) <- c("pred")
price.pred.Conv$actual <- avocadoData2018$AveragePrice
price.pred.Conv$diff <- abs(price.pred.Conv$pred - price.pred.Conv$actual)
mean(price.pred.Conv$diff)

#### Predictions for Organic
## pm.Org.3
price.pred.Org <- data.frame(predict(pm.Org.4, avocadoData2018))
names(price.pred.Org) <- c("pred")
price.pred.Org$actual <- avocadoData2018$AveragePrice
price.pred.Org$diff <- abs(price.pred.Org$pred - price.pred.Org$actual)
mean(price.pred.Org$diff)

```


SCRAP WORK/STUFF I DIDNT WANT TO DELETE BUT DUNNO WHAT TO DO WITH
```{r}
g2 <- lm(AveragePrice ~ DateNumeric + I(DateNumeric ^2), data = AvocadoRegionsOrg)

summary(g2)
#####summary(g)


ggplot(AvocadoRegionsOrg, aes(DateNumeric, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 2, raw = TRUE))

ggplot(AvocadoRegionsOrg, aes(DateNumeric, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 3, raw = TRUE))

ggplot(AvocadoRegionsOrg, aes(DateNumeric, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 4, raw = TRUE))

ggplot(AvocadoRegionsOrg, aes(Total.Volume, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 5, raw = TRUE))

ggplot(AvocadoRegionsOrg, aes(Total.Volume, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 6, raw = TRUE))

g3 <- lm(AveragePrice ~ DateNumeric + I(DateNumeric ^2), data = AvocadoRegionsOrg)

summary(g3)

#######

ggplot(AvocadoRegionsOrg, aes(DateNumeric, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 6, raw = TRUE))

g6 <- lm(AveragePrice ~ DateNumeric + I(DateNumeric ^6), data = AvocadoRegionsOrg)
g66 <- lm(AveragePrice ~ poly(DateNumeric, 8, raw = TRUE), data = AvocadoRegionsOrg)

summary(g66)
summary(g6)


ggplot(AvocadoRegionsOrg, aes(DateNumeric, AveragePrice) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 7, raw = TRUE))

g7 <- lm(AveragePrice ~ DateNumeric + I(DateNumeric ^7), data = AvocadoRegionsOrg)

summary(g7)
```




Things to do:
Filter out the total US data 
Data Cleaning: filter between regions and cities
Filter by individual regions for data viewing purposes 
Need to discuss variance 
Run multiple regressions and compare them using an Anova test




Things to complete by Monday:
- Figure out how to use Polynomial regression correctly
- Predicting/"forecasting" the average price of avocados give the variables 
- Map a regress onto our data
- ANOVA to compare linear and different degree polynomials 
- Standardizing our R markdown document for group dispersal. 
- Consistently mirror conventional and organic avocado data throughout the presentation/paper
- Comparing two variables to test dependency 
- Preliminary slide deck/power point started














